\thispagestyle{empty} 
\leavevmode%
\vspace*{\fill}\noindent
\begin{center}
  \textbf{DECLARATION OF AUTHORSHIP}
\end{center}

I, Julian Alexander Gercke, declare that this thesis titled, `Supporting Explainable AI on Semantic Constraint Validation' and the work presented in it are my own. I confirm that:

\begin{itemize} 
\item[\tiny{$\blacksquare$}] This work was done wholly or mainly while in candidature for a research degree at this University.
 
\item[\tiny{$\blacksquare$}] Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
 
\item[\tiny{$\blacksquare$}] Where I have consulted the published work of others, this is always clearly attributed.
 
\item[\tiny{$\blacksquare$}] Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.
 
\item[\tiny{$\blacksquare$}] I have acknowledged all main sources of help.
 
\item[\tiny{$\blacksquare$}] Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.
\\
\end{itemize}
\vspace{1cm}
\newlength\us
\settowidth{\us}{-~~~~~~Julian Alexander Gercke~~~~~~-}
\begin{minipage}{0.5\linewidth}
    \centering
    \begin{tabular}{p{\us}}\hline
    \centering\footnotesize Place, Date
    \end{tabular}
\end{minipage}
\begin{minipage}{0.5\linewidth}
    \centering
    \begin{tabular}{p{\us}}\hline
    \centering\footnotesize Julian Alexander Gercke
    \end{tabular}
\end{minipage}


\hfill\strut
\vspace*{\fill}
